### 2018-04-24 contents/reviews/pain.tex
Updates from ShareLaTeX

Another spinothalamic termination site is the mediodorsal (MD) nucleus of the thalamus, where pain afferents of the spinothalamic tract that originated from dorsal horn lamina I layer synapse. The MD nucleus projects third order afferents to the anterior/middle cingulate cortex (ACC/MCC), which is shown to be prominent in the emotive/suffering cognitive aspect of pain.

@@

\paragraph{Sensory-Discrimination}
Cortical regions associated with sensory-discriminative aspect of pain, such as the location and intensity of pain, include the S1 ana S2 cortices. The S1 cortex is responsible for the discrimination of fine-touch sensations. S1 receives inputs directly from VP thalamus. Nonetheless, S1 has not been consistently reported in acute pain studies, possibly due to the difficulty in detecting pain-related activations as there are relatively small number of nociceptive neurons within S1 \cite{bushnell1999pain}. S2, however, receives direct nociceptive input from VPI thalamus, and its BOLD activation is more consistently observed across pain fMRI studies \cite{Wager2013,Brodersen2012}. The posterior insula also receives direct nociceptive inputs from the VMpo thalamus \cite{Craig2003}, and shows strong functional and structural connectivity with S2 \cite{Wiech2014}. Therefore the insula, as well as S1, are believed to plays an important role in sensory-discrimination aspect of pain.

@@

\paragraph{Default mode network}
The default mode network (DMN) are regions of the brain, including the cingulate cortex, which show highly correlative BOLD activations in fMRI during wakeful rest \cite{Greicius2003,Raichle2015}. The core regions associated with the default network are ventral medial prefrontal cortex (vmPFC), posterior cingulate, inferior parietal, lateral temporal cortex, dorsal medial prefrontal cortex (dMPFC) and hippocampal formation. Regions such as vmPFC, dMPFC and the anterior cingulate overlap with the affective-motivational pain matrix, and are found to be disrupted in DMN in chronic pain \cite{Baliki2008}. Pain related DMN deactivation have been found to be attenuated  in mind wandering away from pain, and may reflect antinociceptive processes \cite{Kucyi2013}, and that the DMN may undergo functional reorganization under chronic pain \cite{Baliki2014}.

@@

The insula is an important neuroanatomical structure that has been implicated in both acute \cite{Lorenz2005,Starr2009} and chronic pain \cite{May2008,Jensen2016}. Anatomically it is subdivided into anterior and posterior components by the central insular sulcus. It can be further identified by its gyral features, consisting of 5 major gyri: anterior, middle and posterior short gyri in the anterior insula, and anterior long, posterior long gyri of the posterior insula. These gyri are not found consistently in the general population, but rather show individual variability \cite{Rosen2015}. Craig postulates that the insula is a key region for homeostasis and interoception \cite{Craig2003b}, and is necessary for pain discriminative sensations via VMpo projections in the middle insula in humans \cite{Craig2003a}.

@@

Neuropathic pain can also arise from changes in the central neuroaxis. Trauma and other disruptions to the spinal pathways can affect noxious signal transmission. White matter pathways within the brain may also become damaged by conditions such as trauma, infections, and multiple sclerosis. It is believed that injury to white matter fibers can affect action potential conduction between cortical and subcortical structures. This may give rise or contribute to neuropathic pain \cite{Geha2008}.

>>>>

Another spinothalamic termination site is the mediodorsal (MD) nucleus of the thalamus, where pain afferents of the spinothalamic tract that originated from dorsal horn lamina I layer synapse. The MD nucleus projects third order afferents to the anterior/middle cingulate cortex (ACC/MCC), which is shown to be prominent in the emotive/suffering cognitive aspect of pain \cite{Vogt2016}.

@@

\paragraph{Sensory-Discrimination}
Cortical regions associated with sensory-discriminative aspect of pain, such as the location and intensity of pain, include the S1 and S2 cortices. The S1 cortex is responsible for the discrimination of fine-touch sensations. S1 receives inputs directly from VP thalamus. Nonetheless, S1 has not been consistently reported in acute pain studies, possibly due to the difficulty in detecting pain-related activations as there are relatively small number of nociceptive neurons within S1 \cite{bushnell1999pain}. S2, however, receives direct nociceptive input from VPI thalamus, and its BOLD activation is more consistently observed across pain fMRI studies \cite{Wager2013,Brodersen2012}. The posterior insula also receives direct nociceptive inputs from the VMpo thalamus \cite{Craig2003}, and shows strong functional and structural connectivity with S2 \cite{Wiech2014}. Therefore the insula, as well as S1, are believed to plays an important role in sensory-discrimination aspect of pain.

@@

\paragraph{Default mode network}
The default mode network (DMN) are regions of the brain, including the cingulate cortex, which show highly correlative BOLD activations in fMRI during wakeful rest \cite{Greicius2003,Raichle2015}. The regions have been shown to consistently decrease in activity during attention-demanding, and non-self-referential tasks,when compared with activities during relaxed non-task states \cite{Raichle2015}. The DMN is believed to instantiates processes that support self-referential activities, emotional processing, and recollection \cite{Raichle2015}. The core regions associated with the default network are ventral medial prefrontal cortex (vmPFC), dorsal medial prefrontal cortex (dMPFC), posterior cingulate, inferior parietal, lateral temporal cortex, and hippocampal formation. Regions such as vmPFC and dMPFC overlap with the affective-motivational pain matrix, and are found to be disrupted in DMN in chronic pain \cite{Baliki2008}. Pain related DMN deactivation have been found to be attenuated  in mind wandering away from pain, and may reflect antinociceptive processes \cite{Kucyi2013}, and that the DMN may undergo functional reorganization under chronic pain \cite{Baliki2014}.

@@

The insula is an important neuroanatomical structure that has been implicated in both acute \cite{Lorenz2005,Starr2009} and chronic pain \cite{May2008,Jensen2016}. Anatomically it is subdivided into anterior and posterior components by the central insular sulcus. It can be further identified by its gyral features, consisting of 5 major gyri: anterior, middle and posterior short gyri in the anterior insula, and anterior long, posterior long gyri of the posterior insula. These gyri are not found consistently in the general population, but rather show individual variability \cite{Rosen2015}. Craig postulates that the insula is a key region for homeostasis and interoception \cite{Craig2003b}, and is necessary for pain discriminative sensations via VMpo projections in the middle insula in humans \cite{Craig2003a}. The anteior insula has been shown by functional connectivity analysis to have interactions with anteior MCC and middle insula with posterior MCC \cite{Taylor2009}. Further suggest that the anterior insula and MCC are involved in an emotional/salience network for interoception and emotional integration and monitoring \cite{Vogt2016}.

@@

Neuropathic pain can also arise from changes in the central neuroaxis. Trauma and other disruptions to the spinal pathways can affect noxious signal transmission. Where repeated and sustained noxious stimuli may cause the central nervous system to adapt and amplify responses to stimulation. These can involve activation-dependence plasticity in neurons of the dorsal horn, which result in windup of action potential discharge of these neurons. Additionally, long-term depression may also result in the inhibitory mechanism in interneurons of the substanita gelatinosa \cite{Woolf2000}. White matter pathways within the brain may also become damaged by conditions such as trauma, infections, and multiple sclerosis. It is believed that injury to white matter fibers can affect action potential conduction between cortical and subcortical structures. This may give rise or contribute to neuropathic pain \cite{Geha2008}.
